---
title: "R Notebook"
output: html_notebook
---

# Visualizing Discrete Variables

```{r}
library(tidyverse)
library(palmerpenguins)
```

```{r}
penguins_df <- penguins %>% drop_na()
penguins_df
```

## Summarizing Discrete Variables

Use `group_by` and `summarize` to summarize data

```{r}
penguins_df %>%
  group_by(species, year) %>%
  summarize(
    n_penguins = n(),
    mean_body_mass = mean(body_mass_g),
    sd_body_mass = sd(body_mass_g)
  )
```

## Factors

Use `pull` and `levels` to see ordering of factors

-   strongly disagree

-   disagree

-   neutral

-   agree

-   strongly agree

```{r}
penguins_df %>%
  pull(island) %>%
  levels()
```

Use `mutate` and `factor(..., levels = c(..))` to change the ordering of factors

```{r}
penguins_df <- penguins_df %>%
  mutate(
    year = factor(year, levels = c(2007, 2008, 2009)),
    species = factor(species, levels = c("Adelie", "Gentoo", "Chinstrap"))
  )

penguins_df
```

## Bar Charts

`geom_bar(..., stat = "identity")`

```{r}
penguins_df %>%
  group_by(species) %>%
  summarize(
    n_penguins = n(),
    mean_body_mass = mean(body_mass_g),
    sd_body_mass = sd(body_mass_g)
  ) %>%
  ggplot() +
    geom_bar(
      mapping = aes(y = species, x = n_penguins),
      stat = "identity"
    )
```

Add `fill` and `position = dodge` to make grouped bar chart

```{r}
penguins_df %>%
  group_by(species, year) %>%
  summarize(
    n_penguins = n(),
    mean_body_mass = mean(body_mass_g),
    sd_body_mass = sd(body_mass_g)
  ) %>%
  ggplot() +
    geom_bar(
      mapping = aes(fill = species, y = n_penguins, x = year),
      stat = "identity",
      position = "dodge"
    )
```

## Pie Chart

<https://r-graph-gallery.com/piechart-ggplot2.html>

```{r}
penguins_df %>%
  group_by(species) %>%
  summarize(
    n_penguins = n(),
    mean_body_mass = mean(body_mass_g)
  )
```

```{r}
penguins_df %>%
  group_by(species) %>%
  summarize(
    n_penguins = n(),
    mean_body_mass = mean(body_mass_g)
  ) %>%
  ggplot() +
    geom_bar(
      mapping = aes(x = " ", fill = species, y = n_penguins),
      stat = "identity"
    ) +
    coord_polar("y", start=0) +
    theme_void()
```
